I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-25 Paper/Paper_noProof.tex
table 6
edits

@@

Our experimental evaluations have 3 goals:
\begin{enumerate}
	\item Evaluate experimentally how the choice of the set $Z$ impacts the accuracy of $\MILP_Z$, and compare different choices for $K$ the size of $Z$,
	\item Evaluate on benchmarks the percentage of images that can be verified
	using compensating strength, and compare with other value abstraction verifiers,
	\item Evaluate how the runtime of compensate($K$) scale with the size of DNNs.
\end{enumerate}

\subsection{Setting}
The benchmarks were run on an AMD 5975WX ($32$ cores$@3.6$GHz, 7nm, year 2022) with 256 GB of main memory and 2 NVIDIA RTX 3090. Gurobi 9.52 was used for solving MILP and LP problems. Our implementations for Compensate($K$) is entirely realized in Python 
$3.8.10$, and therefore not particularly optimal.

@@

@@

		\hline
		\text{DNN}  & DeepPoly & PRIMA &  \multicolumn{2}{@{}c@{}|}{\text{ $\alpha$-$\beta$-CROWN (pureBaB) \,}} & 
		{\color{blue}Comp (adapt.)} & Upper\\ 
		& / CROWN & (refined) & TO=30s & TO=2000s & {\color{blue}fast} & Bound\\
		\hline \hline
		$6\times500$\  &   $26\%$ & $32.5\%$ & $41\%$ & $44.5\%$  & {\color{blue}$\mathbf{65\%}$} & $92\%$ \\ 
		$\epsilon = 0.035$ & 34s & 119s & 18.4s & 954s & {\color{blue}925s} &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}
	\vspace{-1cm}
\end{table}

\vspace{-0.4cm}
\begin{figure}[h!]
	\centering
\begin{tikzpicture}
	\begin{axis}[width=10cm,height=5cm,
		xlabel={},
		ylabel={},
		legend pos=south east,
		ymin=0,
		xtick={1,2,3,4,5}, 
		xticklabels={$5\times 100$, $8\times 100$, $5\times 200$, $8\times 200$, $6\times 500$},
		]

@@

	\end{axis}
\end{tikzpicture}
\caption{runtime (seconds per node)}
\label{fig2}
\end{figure}
\vspace{-0.4cm}

@@

\begin{table}[]
	\centering
	\begin{tabular}{|c|c|cc||cc||c|}
		\hline
		\text{DNN}  & DeepPoly &  \multicolumn{2}{@{}c@{}|}{\text{ $\alpha$-$\beta$-CROWN (refinedBaB) \,}} & 
		\multicolumn{2}{@{}c@{}|}{\text{ \color{blue}\CMP slow \,}}
		& Upper\\ 
		& / CROWN & TO=300s & TO=3000s & {\color{blue}adapt} & {\color{blue}fixed} & Bound\\
		\hline \hline
		$8\times200$\  &   $25\%$? & $72\%$? & $\mathbf{79\%}$ & {\color{blue}$75.5\%$}  & {\color{blue}$77\%$} & $91\%$ \\ 
		$\epsilon = 0.015$ & 17s & 95s? & 442s & {\color{blue}391s} & {\color{blue}562s} &\\ \hline
	\end{tabular}
	\caption{WIP! $\%$ of verified images over 200 images. To see how refined BAB scale with TimeOut.}
	\label{tab:example4}
\end{table}

>>>>

\newcommand{\toolname}{\ensuremath{\mathsf{Comp}}

@@

We implemented a prototype of {\toolname} in Python 3.8.  We conducted our evaluation on an AMD 5975WX  ($32$ cores$@3.6$GHz, 7nm, year 2022) with 256 GB of main memory and 2 NVIDIA RTX 3090. Gurobi 9.52 was used for solving MILP and LP problems.

@@

\begin{enumerate}
	\item Evaluate experimentally how the choice of the set $Z$ impacts the accuracy of $\MILP_Z$, and compare different choices for $K$ the size of $Z$,
	\item Evaluate on benchmarks the percentage of images that can be verified
	using compensating strength, and compare with other value abstraction verifiers,
	\item Evaluate how the runtime of compensate($K$) scale with the size of DNNs.
\end{enumerate}

@@

		\hline
		\text{DNN}  & DeepPoly & PRIMA &  \multicolumn{2}{@{}c@{}|}{\text{ $\alpha$-$\beta$-CROWN (pureBaB) \,}} & 
		{\color{blue}Comp (adapt.)} & Upper\\ 
		& / CROWN & (refined) & TO=30s & TO=2000s & {\color{blue}fast} & Bound\\
		\hline \hline
		$6\times500$\  &   $26\%$ & $32.5\%$ & $41\%$ & $44.5\%$  & {\color{blue}$\mathbf{65\%}$} & $92\%$ \\ 
		$\epsilon = 0.035$ & 34s & 119s & 18.4s & 954s & {\color{blue}925s} &\\ \hline
	\end{tabular}
	\caption{$\%$ of verified images and average runtime in seconds, over 200 images.}
	\label{tab:example3}

\end{table}

\begin{figure}[h!]
	\centering
\begin{tikzpicture}
	\begin{axis}[width=10cm,height=5cm,
		xlabel={},
		ylabel={},
		legend pos=south east,
		ymin=0,
		xtick={1,2,3,4,5}, 
		xticklabels={$5\times 100$, $8\times 100$, $5\times 200$, $8\times 200$, $6\times 500$},
		]

@@

	\end{axis}
\end{tikzpicture}
\caption{runtime (seconds per node)}
\label{fig2}
\end{figure}

@@

\begin{table}[]
	\centering
	\begin{tabular}{|c|c|cc||cc||c|}
		\hline
		\text{DNN}  & DeepPoly &  \multicolumn{2}{@{}c@{}|}{\text{ $\alpha$-$\beta$-CROWN (refinedBaB) \,}} & 
		\multicolumn{2}{@{}c@{}|}{\text{ \color{blue}\CMP slow \,}}
		& Upper\\ 
		& / CROWN & TO=300s & TO=3000s & {\color{blue}adapt} & {\color{blue}fixed} & Bound\\
		\hline \hline
		$8\times200$\  &   $25\%$? & $72\%$? & $\mathbf{79.5\%}$ & {\color{blue}$75.5\%$}  & {\color{blue}$77\%$} & $91\%$ \\ 
		$\epsilon = 0.015$ & 17s & 95s? & 451s & {\color{blue}391s} & {\color{blue}562s} &\\ \hline
	\end{tabular}
	\caption{WIP! $\%$ of verified images over 200 images. To see how refined BAB scale with TimeOut.}
	\label{tab:example4}
\end{table}

### 2024-01-25 Paper/introduction.tex

Our three main contributions address the challenges of inaccuracies in neural network computations due to compensating pairs of paths:
\begin{enumerate}
	\item  Our first main contribution establishes, through Theorem \ref{th1}, the foundational understanding that the compensating pairs are the 	%primary 
	source of {\color{red}all} inaccuracies, 
	{\color{red} as in their absence, even the simplest Box abstraction (and most abstractions under consideration) is fully accurate.}
	This revelation, while theoretically pivotal, has limited practical applicability as most networks inherently possess compensating pairs. Nonetheless, this insight lays the groundwork for our innovative approach, {\color{red}{\em relaxing} the usual but inefficient Mixed Integer Linear Program (MILP) formulation encoding exactly DNNs \cite{MILP}. To compute bounds for $z$, we encode as binary variables (the costly ones) only ReLU nodes on compensating paths {\em with target $z$}; other nodes, in particular those on compensating paths to neurons before $z$ are encoded using the much more efficient linear relaxation. To take into account indirectly the compensating paths to previous nodes, we inductively compute bounds for every neuron layer by layer, and use these bounds of previous neurons in the MILP encoding for $z$. This is more efficient, as the number of binary nodes is strongly reduced. However, the question remains how accurate is it?}

\item {\color{red} Our second main contribution, delineated in Theorem \ref{th2}, answers this question, showing that this abstraction is fully accurate, in that the bounds computed in such a way for neuron $z$ correspond exactly to the mininmal and maximal values reachable by inputs of the DNN. Again, the prevalence of compensating paths in actual DNNs renders this algorithm computationally intensive because the number of binary variables is still too high.}

\item Our third main contribution addresses this practical concern, introducing 
\CMP (Algorithm \ref{algo1}), a pragmatic solution that leverages our theoretical insights. This approach ranks ReLU nodes based on {\color{red} 
the weights of the compensating paths they belong to,
and differentiates their treatment in the MILP model (binary for heavy, linear for light). 
%Notably, Algorithm \ref{algo1} is exceptionally amenable to parallelization, as all nodes within the same layer can be processed simultaneously. Furthermore, 
Our comprehensive empirical evaluation indicates that \CMP provides favorable trade-offs between accuracy and completeness compared to existing methods, 
%lowering the number of 
%undecided inputs from $17.6\%$ ($\beta$-Crown \cite{crown}) to $10.4\%$ 
%on DNN $8 \times 200$ (Table \ref{tab:example}), even 
verifying more than $20\%$ more images than $\alpha,\beta$-CROWN \cite{xu2020fast} (or PRIMA \cite{prima}) on the larger $6 \times 500$ DNN trained naturally (Table \ref{tab:example3}).}

>>>>

Our three main contributions address the challenges of inaccuracies in neural network computations due to compensating pairs of paths:
\begin{enumerate}
	\item  Our first main contribution establishes, through Theorem \ref{th1}, the foundational understanding that the compensating pairs are the 	%primary 
	source of all inaccuracies, 
	 as in their absence, even the simplest Box abstraction (and most abstractions under consideration) is fully accurate.
	This revelation, while theoretically pivotal, has limited practical applicability as most networks inherently possess compensating pairs. Nonetheless, this insight lays the groundwork for our innovative approach, {\em relaxing} the usual but inefficient Mixed Integer Linear Program (MILP) formulation encoding exactly DNNs \cite{MILP}. To compute bounds for $z$, we encode as binary variables (the costly ones) only ReLU nodes on compensating paths {\em with target $z$}; other nodes, in particular those on compensating paths to neurons before $z$ are encoded using the much more efficient linear relaxation. To take into account indirectly the compensating paths to previous nodes, we inductively compute bounds for every neuron layer by layer, and use these bounds of previous neurons in the MILP encoding for $z$. This is more efficient, as the number of binary nodes is strongly reduced. However, the question remains how accurate is it?

\item  Our second main contribution, delineated in Theorem \ref{th2}, answers this question, showing that this abstraction is fully accurate, in that the bounds computed in such a way for neuron $z$ correspond exactly to the mininmal and maximal values reachable by inputs of the DNN. Again, the prevalence of compensating paths in actual DNNs renders this algorithm computationally intensive because the number of binary variables is still too high.

\item Our third primary contribution addresses a practical concern by introducing Algorithm \ref{algo1}, a pragmatic solution that builds on our theoretical insights. This method prioritizes ReLU nodes based on the weights of their associated compensating paths and differentiates their treatment in the MILP model, employing binary constraints for heavier nodes and linear constraints for lighter ones. Our extensive empirical evaluation reveals that Algorithm \ref{algo1} achieves a beneficial balance between accuracy and completeness compared to prevailing methods. It reduces the proportion of undecided inputs from 17.6\% (as seen with $\beta$-Crown \cite{crown}) to 10.4\% on the DNN architecture of $8 \times 200$ (refer to Table \ref{tab:example}). Moreover, it verifies over 20\% more images than $\alpha,\beta$-CROWN \cite{xu2020fast} (or PRIMA \cite{prima}) on the larger $6 \times 500$ DNN configuration (as detailed in Table \ref{tab:example3}).
